Bayesian Sparsification of Deep $\mathbb{C}$-valued Networks

### 2020-06-23 paper/paper.tex
dropped appendix about linearops

\title{Bayesian Sparsification Methods for Deep Complex-valued Networks}
\author{Ivan Nazarov, and Evgeny Burnaev}

@@

%% drafting macro
\newcommand{\important}[1]{\textbf{\!\colorbox{red}{#1}\!}}
\newcommand{\attn}[2]{\textbf{\color{red} #2~\textsuperscript{\textit{[#1]}}}}
\newcommand{\verify}[1]{\textit{\!\colorbox{red}{#1}\!}}
\newcommand{\rewrite}[1]{\attn{rewrite}{#1}}
\newcommand{\todo}[1]{{\color{blue} [TODO]} \important{#1}}

%% red-highlight missing citations
\usepackage{etoolbox}
\makeatletter
\patchcmd{\@citex}{\bfseries ?}{\colorbox{red}{\bfseries ?}}{}{}
\makeatother

\begin{document}
\maketitle

@@

Variational Dropout uses \emph{fully factorized Gaussian} approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\omega_j \vert\, \mu_j, \alpha_j \mu_j^2)
$ and factorized scale invariant log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. \citet{molchanov_variational_2017} noticed that $\alpha_j$ reflects the relevance of the
parameter $\omega_j$ it is associated to by being the ratio of its squared mean to its
effective variance. Based on this observation they proposed \emph{Sparse Variational Dropout},
a modification that enables automatic model sparsification by optimizing $\alpha_j$ for each
individual parameter. \citet{louizos_bayesian_2017} extended the idea to structured sparsity
by considering hierarchical prior and variational approximation. They grouped the parameters
$\omega_j$ and coupled them within each one through a shared latent variable, which on the
whole enabled pruning entire input features in each layer.
% {louizos_bayesian_2017} eq. (11) variance -> gaussian entropy -> informative bits
% They also proposed to estiamte bit precision for quantization based on parameter's
% estimated uncertainty (variance or entropy).
%
Due to factorization assumption, the term $KL(q_\theta \| \pi)$ in \eqref{eq:sgvb_estimator}
for Sparse VD unravels into $
  \sum_j K(\tfrac{\sigma^2_{j}}{\mu_{j}^2})
$ with
\begin{equation}  \label{eq:improper-kl-div-real}
  K(\alpha)
    \propto \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximated $K(\alpha)$ over $\alpha \in (0, 1)$ by a
polynomial with a logarithmic term, and later \citet{molchanov_variational_2017} refined
the approximation of \eqref{eq:improper-kl-div-real} by weighted sum of a sigmoid and a
soft-plus term.
% see p.6 sec 3.4 of kingma_variational_2015
In appendix~\ref{sec:real-chisq-grad} we verify the derivative of their approximation
against a Monte-Carlo estimate for $\alpha$ varying over a fine $\log$-scale grid and
the exact expression for gradient of \eqref{eq:improper-kl-div-real}.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates
% (in) it is in poor taste to brag about this beign the first time anyone has derived a
% derivative of an intractable function.

@@

This $\cplx \simeq \real^2$ identification allows straightforward retrofitting of $\cplx$VNN
into existing $\real$-valued auto-differentiation frameworks for deep learning. This
act is backed by Wirtinger ($\cplx\real$) calculus, which enables generalized treatment
of functions of complex argument, by regarding $z$ and its \emph{complex conjugate}
$\conj{z}$ as independent variables and defining derivative operators with respect to
them through partial derivatives with respect to real and imaginary parts. These definitions
simplify manual analysis of $\cplx$ derivatives and satisfy the product and chain rules, respect
complex conjugation and linearity for $\cplx \to \cplx$ maps, and such were used to define
$\cplx$ version of back-propagation, \citep{benvenuto_complex_1992,guberman_complex_2016}.
However, since auto-differentiation frameworks can algorithmically handle computational
graphs of arbitrary complexity, explicit use of Wirtinger derivatives is not required,
especially considering the fact that the direction of the steepest ascent of a $
  \cplx \to \real
$ function is given by complex conjugate gradient $\nabla_{\conj{z}}$, which coincides
with the classical gradient of the same function viewed as $\real^2 \to \real$,
(see appendix~\ref{sub:wirtinger_calculus}).
%
% in natively $\cplx$-valued backprop it would be necessary to modify the existing sgd
% to explicitly use $\nabla_{\conj{z}} f(z)$, while no such modification is necessary
% in re-im. If some inner activation of a $\cplx$VNN is holomorphic, i.e. satisfies
% Cauchy-Riemann conditions, then that particular step of in the chain rule that is
% backprop can be simplified by assuming one piece of the tracked gradient to be exact
% zero. But other wise it is necessary to track both $\partial_{\conj{z}}$ and $\partial_{z}$.
%
% Training a $\cplx$VNN involves optimizing a real-valued objective with respect to
% complex valued arguments.
%
% The computed gradients are theoretically justified
% - CR dictates the it is necessary to track both $z$ and $\conj{z}$ derivatives
% - the direction of largest increase in f: C-R is given by the gradient with
% respect to the conjugate, which coincides in the re-im representation with 
% the ordinary gradient wrt u and v
% - the Chain Rule in CR holds and when a holomorphic C-C function is backproped
% the autodiff just wastes cycles on computing its conj-derivative (rather than assuming
% it to be zero)

@@

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (VD, \texttt{fft}, MNIST):
    $2 \real / \cplx$ (\textit{top}) and $\real / \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__fft}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (ARD, \texttt{raw}, MNIST):
    $\real / \cplx$ (\textit{top}) and $\real / \tfrac12 \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__raw}
\end{figure}

@@

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    % \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
    \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy profile for the $\real$ and $\cplx$ VGG16.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__musicnet__trade-off/paper__musicnetram__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Performance-compression curve for \textbf{\color{blue} VD}, \textbf{\color{orange} ARD},
    and the ${\color{gray} k_3}$ version compressed with VD.
  }
  \label{fig:musicnet__trade-off}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__fine-tune_fx__early.pdf}
  \end{subfigure}
  \caption{%
    Early stopping epoch at fine-tuning stage.
  }
  \label{fig:musicnet__early_stopping}
\end{figure}

@@

In this appendix we study the approximation proposed by \citet{molchanov_variational_2017}
for the KL divergence term \eqref{eq:improper-kl-div-real}. We show empirically that
the gradient feedback provided by the approximation is sufficiently close to $
  \tfrac{d}{d\log \alpha} K(\alpha)
$, making it a viable alternative to exact gradients, computed in \citet{hron_variational_2018}.
% The derivation in this section different from \citet{lapidoth_capacity_2003}.
% The results are identical: they use DCT to establish interchangeability of limits
% (sum-integral-derivative), whilst we allude to calculus theorems on differentiation
% and integration of power series within the radius of their convergence (which are
% essentially corollaries to the DCT).
$. The divergence \eqref{eq:improper-kl-div-real} has the form
distributions: if $Z_{\mid J} \sim \chi^2_{k + 2J}$ for $
  \mathbb{E}_{W\sim \chi^2_k(\lambda)} \log W
Since $\chi^2_\nu$ is Gamma distribution $\Gamma(\tfrac\nu2, \tfrac12)$, its logarithmic
moment $
$, where $\psi(x)$ is the digamma function. Hence
  \mathbb{E}_{W\sim \chi^2_k(\lambda)} \log W
    = \log2 + \mathbb{E}_{J\sim {\mathcal{P}ois}(\tfrac\lambda2)}
    % = \log2
    % + \sum_{j\geq 0}
    %     e^{- \tfrac{\lambda}2}
    %     \frac{\bigl(\tfrac{\lambda}2\bigr)^j}{j!}
    %   \psi\bigl(\tfrac{m+2J}{2}\bigr)
We can use REINFORCE method to compute the derivative of \eqref{eq:exp_digamma} with respect
to $\lambda$ \citep{williams_simple_1992}. Since Poisson $\log$-probability is $
  \log p_\lambda(J)
    = - \tfrac{\lambda}2
      + J \log \tfrac{\lambda}2
      - \log J!
$, we get
\begin{equation}  \label{eq:exp_reinforce}
  \frac{d \eqref{eq:exp_digamma}}{d \lambda}
    =
    \mathbb{E}_{J\sim {\mathcal{P}ois}(\tfrac\lambda2)}
      \tfrac{d}{d \lambda} \log p_\lambda(J)
    % = \mathbb{E}_{J}
    %   \bigl(
    %     \tfrac2\lambda J - 1
    %   \bigr)
    %   \frac{\psi\bigl(\tfrac{m + 2 J}{2}\bigr)}2
    = \frac12 g_m\bigl(\tfrac\lambda2\bigr)
where
    =
    \mathbb{E}_{J\sim {\mathcal{P}ois}(x)}
      \bigl( \tfrac{J}x - 1 \bigr)
    =
    e^{- x} \sum_{j \geq 0}
      \frac{x^j}{j!} \bigl( \tfrac{J}x - 1 \bigr)
By manipulating the partial sums within $g_m(x)$ and making use of the property $
  \psi(x + 1) = \psi(x) + \tfrac1x
$ of the digamma funciton for $x > 0$, we get
      \frac1{j!}
      \frac{x^{j + \tfrac{m}2}}{j + \tfrac{m}2}
Since for any $J$ the functions $
$ are positive on $(0, x)$ and upper bounded by integrable $
$, which turns out to be the limit of the partial sums, the Dominated Convergence
theorem implies
      \int_0^x e^t t^{(\tfrac{m}2 - 1)} dt
  F_m
  \frac{d \eqref{eq:exp_digamma}}{d \lambda}
    = \frac12 g_m\bigl(\tfrac\lambda2\bigr)
    = \Bigl(\sqrt{\tfrac2\lambda}\Bigr)^m F\Bigl(\sqrt{\tfrac\lambda2}\Bigr)
    = - \frac1{\sqrt{2\alpha}} F\bigl(\tfrac1{\sqrt{2\alpha}}\bigr)

We compute the Monte-Carlo estimator of \eqref{eq:improper-kl-div-real} on a sample of $10^7$
draws over an equally spaced grid of $\log \alpha$ in $[-12, +12]$ of size $4096$. The
coefficients of the approximation, proposed by \citet{molchanov_variational_2017}, are
within $1\%$ relative tolerance of the reported therein. An equivalent expression of the
approximation is given in \eqref{eq:improper-kl-div-real-approx}.
% see eq.~(14): flit overall sign, use $C = -k_1$, $1-\sigma(x) = \sigma(-x)$
\begin{equation}  \label{eq:improper-kl-div-real-approx}
  \eqref{eq:improper-kl-div-real}
% - KL(\mathcal{N}(w \mid \mu, \alpha \mu^2) \|
%     \tfrac1{\lvert w \rvert})
%   \approx
%     k_1 \sigma(k_2 + k_3 \log \alpha)
%     + C \big\vert_{C = -k1}
%     - \tfrac12 \log (1 + e^{-\log \alpha})
  % = \tfrac12 \log \alpha
  %   - \mathbb{E}_{\varepsilon \sim \mathcal{N}(1, \alpha)}
  %   \log{\lvert \varepsilon \rvert} + C
  \approx
    \frac12 \log{\bigl(1 + e^{-\log \alpha}\bigr)}
    + k_1 \sigma\bigl(- (k_2 + k_3 \log \alpha)\bigr)
    % - k_1 \bigl(1 - \sigma(k_2 + k_3 \log \alpha)\bigr)
  \,,
\end{equation}
% in parctial implementations we suggest using softplus to numerically comptue the last
% term, since it is typically implemented in a floating-point stable fashion.
The numerically estimated derivative of the penalty term with respect to $\log \alpha$ using
forward differences seems to be following \eqref{eq:real-kl-div-deriv-log} very closely (up to
sampling error), see fig.~\ref{fig:molchanov-derivative-replica}. The derivative of
\eqref{eq:improper-kl-div-real-approx} with respect to $\log \alpha$ appears to be very
close to \eqref{eq:real-kl-div-deriv-log}. For sake of completeness, we compute a similar
Monte-Carlo estimate for $\cplx$ variational dropout KL divergence term in \eqref{eq:c-vd-kl-div}
with $\beta = 2$, fit the best approximation \eqref{eq:improper-kl-div-real-approx}, and
compare it against the exact $\log \alpha$ derivative $
  \frac{d \eqref{eq:c-vd-kl-div}}{d\log \alpha}
    % = \frac{d \alpha}{d\log \alpha} \frac{d}{d\alpha} \bigl(
    %   \log \tfrac1\alpha - \mathop{Ei}\bigl(-\tfrac1\alpha\bigr)
    % = \alpha \bigl(
    %   - \tfrac1\alpha
    %   - \tfrac{e^x}{x} \bigg\vert_{x=-\tfrac1\alpha} \alpha^{-2}
    % \bigr)
    = e^{-\tfrac1\alpha} - 1
$.

@@

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__raw}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__raw}
\end{figure}

>>>>

\usepackage{microtype}
\usepackage{hyperref}

\usepackage[accepted]{icml2020/icml2020}

% \title{Bayesian Sparsification Methods for Deep Complex-valued Networks}
% \author{Ivan Nazarov, and Evgeny Burnaev}

\icmltitlerunning{Bayesian Sparsification of Deep $\mathbb{C}$-valued networks}

@@

\begin{document}
\twocolumn[
\icmltitle{Bayesian Sparsification of Deep $\mathbb{C}$-valued Networks}

\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
  \icmlauthor{Ivan Nazarov}{sk}
  \icmlauthor{Evgeny Burnaev}{sk}
\end{icmlauthorlist}

\icmlaffiliation{sk}{
  Centre for Data Intensive Sciecne and Engineering,
  Skolkovo Insitiute of Science and Technology, Msocow, Russia
}
\icmlcorrespondingauthor{Ivan Nazarov}{ivan.nazarov@skolkovotech.ru}

% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{complex-valued neural networks, Variational Dropout, network compression}

\vskip 0.3in
]
\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% \printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

@@

Variational Dropout uses \emph{fully factorized Gaussian} approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\omega_j \vert\, \mu_j, \alpha_j \mu_j^2)
$ and factorized scale invariant log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. \citet{molchanov_variational_2017} noticed that $\alpha_j$ reflects the relevance of the
parameter $\omega_j$ it is associated to by being the ratio of its squared mean to its
effective variance. Based on this observation they proposed \emph{Sparse Variational Dropout},
a modification that enables automatic model sparsification by optimizing $\alpha_j$ for each
individual parameter. \citet{louizos_bayesian_2017} extended the idea to structured sparsity
by considering hierarchical prior and variational approximation. They grouped the parameters
$\omega_j$ and coupled them within each one through a shared latent variable, which on the
whole enabled pruning entire input features in each layer.
% {louizos_bayesian_2017} eq. (11) variance -> gaussian entropy -> informative bits
% They also proposed to estiamte bit precision for quantization based on parameter's
% estimated uncertainty (variance or entropy).

Due to factorization assumption, the term $KL(q_\theta \| \pi)$ in \eqref{eq:sgvb_estimator}
for Sparse VD unravels into $
  \sum_j K(\tfrac{\sigma^2_{j}}{\mu_{j}^2})
$ with
\begin{equation}  \label{eq:improper-kl-div-real}
  K(\alpha)
    \propto \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\citet{kingma_variational_2015} approximated $K(\alpha)$ over $\alpha \in (0, 1)$ by a
polynomial with a logarithmic term, and later \citet{molchanov_variational_2017} refined
the approximation of \eqref{eq:improper-kl-div-real} by weighted sum of a sigmoid and a
soft-plus term.
% see p.6 sec 3.4 of kingma_variational_2015
In appendix~\ref{sec:real-chisq-grad} we verify the derivative of their approximation
against a Monte-Carlo estimate for $\alpha$ varying over a fine $\log$-scale grid and
the exact expression for gradient of \eqref{eq:improper-kl-div-real}.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates
% (in) it is in poor taste to brag about this beign the first time anyone has derived a
% derivative of an intractable function.

@@

This $\cplx \simeq \real^2$ identification allows straightforward retrofitting of $\cplx$VNN
into existing $\real$-valued auto-differentiation frameworks for deep learning. This
act is backed by Wirtinger ($\cplx\real$) calculus, which enables generalized treatment
of functions of complex argument, by regarding $z$ and its \emph{complex conjugate}
$\conj{z}$ as independent variables and defining derivative operators with respect to
them through partial derivatives with respect to real and imaginary parts. These definitions
simplify manual analysis of $\cplx$ derivatives and satisfy the product and chain rules, respect
complex conjugation and linearity for $\cplx \to \cplx$ maps, and as such were used to define
$\cplx$ version of back-propagation, \citep{benvenuto_complex_1992,guberman_complex_2016}.
However, since auto-differentiation frameworks can algorithmically handle computational
graphs of arbitrary complexity, explicit use of Wirtinger derivatives is not required,
especially considering the fact that the direction of the steepest ascent of a $
  \cplx \to \real
$ function is given by complex conjugate gradient $\nabla_{\conj{z}}$, which coincides
with the classical gradient of the same function viewed as $\real^2 \to \real$,
(see appendix~\ref{sub:wirtinger_calculus}).
%
% in natively $\cplx$-valued backprop it would be necessary to modify the existing sgd
% to explicitly use $\nabla_{\conj{z}} f(z)$, while no such modification is necessary
% in re-im. If some inner activation of a $\cplx$VNN is holomorphic, i.e. satisfies
% Cauchy-Riemann conditions, then that particular step of in the chain rule that is
% backprop can be simplified by assuming one piece of the tracked gradient to be exact
% zero. But other wise it is necessary to track both $\partial_{\conj{z}}$ and $\partial_{z}$.
%
% Training a $\cplx$VNN involves optimizing a real-valued objective with respect to
% complex valued arguments.
%
% The computed gradients are theoretically justified
% - CR dictates the it is necessary to track both $z$ and $\conj{z}$ derivatives
% - the direction of largest increase in f: C-R is given by the gradient with
% respect to the conjugate, which coincides in the re-im representation with 
% the ordinary gradient wrt u and v
% - the Chain Rule in CR holds and when a holomorphic C-C function is backproped
% the autodiff just wastes cycles on computing its conj-derivative (rather than assuming
% it to be zero)

@@

% a couple of plots from MNIST and Fashion-MNIST
\begin{figure}[!t]
  \centering
  \begin{subfigure}  % [b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}  % [b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (VD, \texttt{fft}, MNIST):
    $2 \real / \cplx$ (\textit{top}) and $\real / \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__fft}
\end{figure}

\begin{figure}[!t]
  \centering
  \begin{subfigure}  % [b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}  % [b]{1.\columnwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\columnwidth]{figure__mnist-like__trade-off/appendix__cmp__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy curve (ARD, \texttt{raw}, MNIST):
    $\real / \cplx$ (\textit{top}) and $\real / \tfrac12 \cplx$ (\textit{bottom}).
  }
  \label{fig:mnist-like__trade-off__raw}
\end{figure}

@@

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  % \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
  \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \caption{%
    The compression-accuracy profile for the $\real$ and $\cplx$ VGG16.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=\columnwidth]{figure__musicnet__trade-off/paper__musicnetram__fft__-0.5.pdf}
  \caption{%
    Performance-compression curve for \textbf{\color{blue} VD}, \textbf{\color{orange} ARD},
    and the ${\color{gray} k_3}$ version compressed with VD.
  }
  \label{fig:musicnet__trade-off}
\end{figure}

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=\columnwidth]{figure__fine-tune_fx__early.pdf}
  \caption{%
    Early stopping epoch at fine-tuning stage.
  }
  \label{fig:musicnet__early_stopping}
\end{figure}

@@

@@

\begin{figure}[b]
  \centering
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__fft__-0.5.pdf}
  \end{subfigure} \\ %
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using Fourier features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__fft}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__ARD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of ARD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__ARD__raw}
\end{figure}

\begin{figure}[b]
  \centering
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    % used in main text
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__emnist_letters__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__kmnist__raw__-0.5.pdf}
  \end{subfigure} \\%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}%
  \begin{subfigure}  % [b]{0.5\columnwidth}
    \centering
    \includegraphics[width=0.47\linewidth]{figure__mnist-like__trade-off/appendix__VD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The trade-off of VD method for $\real$ and $\cplx$ models using raw features.
  }
  \label{fig:appendix__mnist-like__trade-off__VD__raw}
\end{figure}
